---
title: "16S rRNA Microbiome Pipeline using DADA2 - full dataset"
subtitle: "Memory-Efficient ASV Inference, Taxonomy, Phyloseq Exploration, and Differential Abundance (Mouse Gut, full dataset)
author: "PDA"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    theme: flatly
    highlight: tango
    code_folding: show
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  eval = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  fig.width = 7,
  fig.height = 5
)
```

Download the 16S reference genome (to be used for taxonomy here: [**https://zenodo.org/records/4587955**](https://zenodo.org/records/4587955){.uri} (SILVA v138.1)) *silva_nr99_v138.1_train_set.fa.gz* - train data *silva_species_assignment_v138.1.fa.gz* - species data

## Package Installation

```{r install-packages}
# CRAN packages
install.packages(c(
  "ggplot2", "devtools", "vegan", "dplyr",
  "tidyverse", "remotes", "ggpubr", "rstatix",
  "stringr", "data.table", "lubridate"
))

if (!requireNamespace("BiocManager", quietly = TRUE))
{
  install.packages("BiocManager")
}
BiocManager::install(c("dada2", "phyloseq", "DESeq2",
                       "ANCOMBC", "Biostrings", "microbiome"))

# nicer plots
devtools::install_github("thomasp85/patchwork")
remotes::install_github("gavinsimpson/ggvegan")
```

## Config paths

```{r config}
library(dada2)
library(data.table)
packageVersion("dada2")

# Paths
# Raw, gzipped FASTQs
data <- "~/mouse_full_data/data"

# project folder for all derived outputs
proj_dir <- "~/mouse_full_data/"

# r object file dir
rds_dir <- file.path(proj_dir, "rds")

# filtered fastqs written by DADA2
filt_dir <- file.path(data, "filtered")

# SILVA training data files(for taxonomy)
ref_dir <- file.path(proj_dir, "reference")

# Directory specifically for saving exported plot graphics
plot_dir <- file.path(proj_dir, "plots")

# Directory specifically saving table outputs (CSVs)
table_dir <- file.path(proj_dir, "tables")

for (d in c(proj_dir, rds_dir, filt_dir, ref_dir, plot_dir, table_dir)) dir.create(d, recursive = TRUE, showWarnings = FALSE)

# SILVA Reference DB
silva_train <- file.path(ref_dir, "silva_nr99_v138.1_train_set.fa.gz")
silva_species <- file.path(ref_dir, "silva_species_assignment_v138.1.fa.gz")

# Threading (for memory and faster processing)
N_THREADS <- max(1, parallel::detectCores() - 1) # for filtering/denoising
TAX_THREADS <- 2 # keep low for taxonomy genx
TAX_CHUNK <- 2000 # ASVs per taxonomy batch

# Helper: report current R memory use
mem_report <- function(tag = "") {
  gc(full = TRUE)
  used_mb <- sum(gc()[, "used"] * c(8,8)) / 1e6
  cat(sprintf("[MEM %s] R heap ~%.0f MB\n", tag, used_mb))
  invisible(NULL)
}
```

## The Dataset

This is the full **Schloss lab** mouse-gut dataset with the following convention

```         
M1D13_S12_L001_R1_001.fastq.gz
└───┘ └─┘ └──┘ └┘ └─┘
  │    │    │   │   └ read chunk index
  │    │    │   └──── read direction (R1 = forward, R2 = reverse)
  │    │    └──────── lane
  │    └───────────── Illumina sample number (S12)
  └────────────────── sample id (M1D13)
```

## Read in and sort the fastq file names

```{r list-fastqs}
# forward (R1) and reverse (R2) reads
dataF <- sort(list.files(data, pattern = "_R1_001\\.fastq\\.gz$", full.names = TRUE))
dataR <- sort(list.files(data, pattern = "_R2_001\\.fastq\\.gz$", full.names = TRUE))

# sample ID
sample.names <- sapply(strsplit(basename(dataF), "_"), `[`, 1)
cat("Forward files:", length(dataF)," Reverse files:", length(dataR), "\n")
cat("First few samples:", paste(head(sample.names), collapse = ", "), "\n")

# sanity checks
stopifnot(length(dataF) == length(dataR))
# confirm R1/R2 line up sample-for-sample
ok <- identical(
  sub("_R1_001\\.fastq\\.gz$", "", basename(dataF)),
  sub("_R2_001\\.fastq\\.gz$", "", basename(dataR))
)
stopifnot(ok)
```

## Step 1: Inspect Read Quality Profiles

```{r qc-forward}
library(ggplot2)
p_qc_f <- plotQualityProfile(dataF[1:16], aggregate = FALSE)
print(p_qc_f)
ggsave(filename = file.path(plot_dir, "qc_forward.png"), plot = p_qc_f, width = 14, height = 10, dpi = 700)
```

```{r qc-reverse}
p_qc_r <- plotQualityProfile(dataR[1:16], aggregate = FALSE) 
print(p_qc_r)
ggsave(filename = file.path(plot_dir, "qc_reverse.png"), plot = p_qc_r, width = 14, height = 10, dpi = 300)
```

## Step 2: Filter and Trim

```{r filter-paths}
filt.dataF <- file.path(filt_dir, paste0(sample.names, "_F_filt.fastq.gz"))
filt.dataR <- file.path(filt_dir, paste0(sample.names, "_R_filt.fastq.gz"))

names(filt.dataF) <- sample.names
names(filt.dataR) <- sample.names
```

| Parameter | Forward | Reverse | Meaning |
|-------------------|------------------|------------------|------------------|
| `truncLen` | 240 | 160 | Truncate reads at this length (guided by quality plots) |
| `maxEE` | 2 | 2 | Max expected errors per read |
| `maxN` | 0 | 0 | No ambiguous bases allowed |
| `truncQ` | 2 | 2 | Truncate at first base with quality ≤ 2 |
| `rm.phix` | TRUE | TRUE | Remove PhiX control reads |

```{r filter-and-trim}
out <- filterAndTrim(
  dataF, filt.dataF, dataR, filt.dataR,
  truncLen = c(240, 160),
  maxN = 0, maxEE = c(2, 2), truncQ = 2, rm.phix = TRUE,
  compress = TRUE, multithread = N_THREADS
)
head(out)

# some samples can lose ALL reads after filtering which can break downstream steps - drop them
kept <- file.exists(filt.dataF) &
  file.exists(filt.dataR)
if (any(!kept)) {
  message("Dropping ", sum(!kept), " sample(s) with no reads after filtering: ",
          paste(sample.names[!kept], collapse = ", "))
}
filt.dataF <- filt.dataF[kept]
filt.dataR <- filt.dataR[kept]
sample.names <- sample.names[kept]

saveRDS(out, file.path(rds_dir, "filter_out.rds"))
saveRDS(sample.names, file.path(rds_dir, "sample_names.rds"))
```

```{r qc-after-filter}
p_post_r <- plotQualityProfile(filt.dataR[1:16], aggregate = FALSE)
ggsave(filename = file.path(plot_dir, "qc_reverse_post_filter.png"), plot = p_post_r, width = 14, height = 10, dpi = 700)

p_post_f <- plotQualityProfile(filt.dataF[1:16], aggregate = FALSE)
ggsave(filename = file.path(plot_dir, "qc_forward_post_filter.png"), plot = p_post_f, width = 14, height = 10, dpi = 700)
```

## Step 3: Learn the Error Rates

```{r learn-errors}
errF <- learnErrors(filt.dataF, multithread = N_THREADS)
errR <- learnErrors(filt.dataR, multithread = N_THREADS)

saveRDS(errF, file.path(rds_dir, "errF.rds"))
saveRDS(errR, file.path(rds_dir, "errR.rds"))
mem_report("after learnErrors")
```

```{r plot-errors}
p_err_f <- plotErrors(errF, nominalQ = TRUE)
ggsave(filename = file.path(plot_dir, "errors_forward.png"), plot = p_err_f, width = 14, height = 10, dpi = 700)

p_err_r <- plotErrors(errR, nominalQ = TRUE)
ggsave(filename = file.path(plot_dir, "errors_reverse.png"), plot = p_err_r, width = 14, height = 10, dpi = 700)
```

## Steps 4 - 5: Denoise + Merge

This is a memory-critical step and requires optimization. Instead of calling `dada2()` on **all** samples at once, loop **one sample at a time**: `dereplicate -> denoise F -> denoise R -> merge -> store only merged result -> free everything else -> gc()`

```{r denoise-merge-loop}
# (Re)load inputs - this block can be rerun in case session ends due to memory depletion 
sample.names <- readRDS(file.path(rds_dir, "sample_names.rds")) 
errF <- readRDS(file.path(rds_dir, "errF.rds")) 
errR <- readRDS(file.path(rds_dir, "errR.rds"))
filt.dataF <- file.path(filt_dir, paste0(sample.names, "_F_filt.fastq.gz")); names(filt.dataF) <- sample.names
filt.dataR <- file.path(filt_dir, paste0(sample.names, "_R_filt.fastq.gz")); names(filt.dataR) <- sample.names

mergers <- vector("list", length(sample.names)); names(mergers) <- sample.names

denoisedF <- integer(length(sample.names)); names(denoisedF) <- sample.names
denoisedR <- integer(length(sample.names)); names(denoisedR) <- sample.names

getN <- function(x) sum(getUniques(x))

for (sam in sample.names) {
  cat("Denoising:", sam, "\n")
  
  derepF <- derepFastq(filt.dataF[[sam]])
  ddF <- dada(derepF, err = errF, multithread = N_THREADS)
  
  derepR <- derepFastq(filt.dataR[[sam]])
  ddR <- dada(derepR, err = errR, multithread = N_THREADS)
  
  merger <- mergePairs(ddF, derepF, ddR, derepR, verbose = FALSE)
  
  mergers[[sam]] <- merger
  denoisedF[[sam]] <- getN(ddF)
  denoisedR[[sam]] <- getN(ddR)
  
  # free per-sample objects before the next iteration
  rm(derepF, derepR, ddF, ddR, merger)
  gc(full = TRUE)
}

saveRDS(mergers, file.path(rds_dir, "mergers.rds"))
saveRDS(denoisedF, file.path(rds_dir, "denoisedF.rds"))
saveRDS(denoisedR, file.path(rds_dir, "denoisedR.rds"))

mem_report("after denoise+merge loop")
```

## Step 6: Make the Sequence Table

```{r make-seqtab}
mergers <- readRDS(file.path(rds_dir, "mergers.rds"))

seqtab <- makeSequenceTable(mergers)
cat("Samples x ASVs:", paste(dim(seqtab), collapse = " x "), "\n")
print(table(nchar(getSequences(seqtab))))

saveRDS(seqtab, file.path(rds_dir, "seqtab.rds"))
```

## Step 7: Remove Chimeras

```{r remove-chimeras}
seqtab <- readRDS(file.path(rds_dir, "seqtab.rds"))

seqtab.nochim <- removeBimeraDenovo(
  seqtab, method = "consensus", multithread = N_THREADS, verbose = TRUE
)

cat("Non-chimeric ASVs:", ncol(seqtab.nochim), "\n")
cat("Fraction of reads kept:", round(sum(seqtab.nochim) / sum(seqtab), 4), "\n")

saveRDS(seqtab.nochim, file.path(rds_dir, "seqtab_nochim.rds"))
```

## Step 8: Track Reads the Pipeline

```{r track-reads}
out <- readRDS(file.path(rds_dir, "filter_out.rds"))
sample.names <- readRDS(file.path(rds_dir, "sample_names.rds"))
denoisedF <- readRDS(file.path(rds_dir, "denoisedF.rds"))
denoisedR <- readRDS(file.path(rds_dir, "denoisedR.rds"))
mergers <- readRDS(file.path(rds_dir, "mergers.rds"))
seqtab.nochim <- readRDS(file.path(rds_dir, "seqtab_nochim.rds"))

getM <- function(x) sum(getUniques(x))

# Align the filter matrix to surviving samples
out2 <- out[paste0(sample.names, "_F_filt.fastq.gz") %in% basename(names(denoisedF)) | TRUE, , drop = FALSE]

# rebuild from saved pieces matched by sample_id
track <- cbind(
  input = out[, "reads.in"][seq_along(sample.names)],
  filtered = out[, "reads.out"][seq_along(sample.names)],
  denoisedF = denoisedF,
  denoisedR = denoisedR,
  merged = sapply(mergers, getN),
  nonchim = rowSums(seqtab.nochim)[sample.names]
)
rownames(track) <- sample.names
head(track, 10)

# Export processing tracking summary table
write.csv(track, file = file.path(table_dir, "pipeline_read_tracking.csv"), row.names = TRUE)
```

## Step 9: Assign Taxonomy - **chunked + low-thread**

```{r assign-taxonomy-chunked}
seqtab.nochim <- readRDS(file.path(rds_dir, "seqtab_nochim.rds"))
stopifnot(file.exists(silva_train)) # check that 16S ref is present

seqs <- getSequences(seqtab.nochim)
chunks <- split(seqs, ceiling(seq_along(seqs) / TAX_CHUNK))
cat("Assigning taxonomy for", length(seqs), "ASVs in", length(chunks), "chunk(s)\n")

taxa_chunks <- vector("list", length(chunks))
for (i in seq_along(chunks)) {
  cat(sprintf(" chunk %d / %d (%d ASVs)\n", i, length(chunks), length(chunks[[i]])))
  taxa_chunks[[i]] <- assignTaxonomy(
    chunks[[i]], silva_train,
    multithread = TAX_THREADS,
    tryRC = TRUE # guards against reverse-complement orientation issues
  )
  gc(full = TRUE) # release the reference + chunk before the next batch
}

taxa <- do.call(rbind, taxa_chunks)
rm(taxa_chunks); gc(full = TRUE)

# Species assignment - only run this if you have the species files
if (file.exists(silva_species)) {
  sp_chunks <- split(rownames(taxa), ceiling(seq_len(nrow(taxa)) / TAX_CHUNK))
  taxa_sp <- vector("list", length(sp_chunks))
  for (i in seq_along(sp_chunks)) {
    sub <- taxa[sp_chunks[[i]], , drop = FALSE]
    taxa_sp[[i]] <- addSpecies(sub, silva_species)
    gc(full = TRUE)
  }
  taxa <- do.call(rbind, taxa_sp)
  rm(taxa_sp); gc(full = TRUE)
}

saveRDS(taxa, file.path(rds_dir, "taxa.rds"))
mem_report("after taxonomy")
```

```{r taxa-inspect}
taxa.print <- taxa
rownames(taxa.print) <- NULL
head(taxa.print)

table(taxa[, "Phylum"], useNA = "ifany")
cat("ASVs with a species-level call:", sum(!is.na(taxa[, "Species"])), "of", nrow(taxa), "\n")
```

## Step 10: Build the Phyloseq object

```{r load-phyloseq}
library(phyloseq)
library(Biostrings)
library(ggplot2)
theme_set(theme_bw())

seqtab.nochim <- readRDS(file.path(rds_dir, "seqtab_nochim.rds"))
taxa <- readRDS(file.path(rds_dir, "taxa.rds"))
```

```{r build-metadata}
samples.out <- rownames(seqtab.nochim)

# group/gender = leading letters; subjects = digits before "D"
gender_code <- sub("^([A-Za-z]+).*$", "\\1", samples.out)
subject <- sub("^[A-Za-z]+([0-9]+)D.*$", "\\1", samples.out)
day <- suppressWarnings(as.integer(sub("^.*D([0-9]+)$", "\\1", samples.out)))

samdf <- data.frame(
  Sample = samples.out,
  Gender = ifelse(gender_code == "F", "Female",
                  ifelse(gender_code == "M", "Male", gender_code)),
  Subject = subject,
  Day = day,
  row.names = samples.out,
  stringsAsFactors = FALSE
)

# Early vs Late
samdf$When <- ifelse(samdf$Day < 100, "Early", "Late")

# Mock/Control samples will not fit longitudinal design
is_ctrl <- grepl("mock|control|blank|ntc", samples.out, ignore.case = TRUE) | is.na(samdf$Day)
samdf$When[is_ctrl] <- NA
samdf$Subject[is_ctrl] <- NA

write.csv(samdf, file = file.path(table_dir, "sample metadata.csv"), row.names = TRUE)
table(samdf$When, useNA = "ifany")
```

## Construct the object

```{r build-phyloseq}
ps <- phyloseq(
  otu_table(seqtab.nochim, taxa_are_rows = FALSE),
  sample_data(samdf),
  tax_table(taxa)
)

# replace long ASV sequences with short labels; keep sequences in refseq()
dna <- Biostrings::DNAStringSet(taxa_names(ps))
names(dna) <- taxa_names(ps)
ps <- merge_phyloseq(ps, dna)
taxa_names(ps) <- paste0("ASV", seq(ntaxa(ps)))

ps
saveRDS(ps, file.path(rds_dir, "phyloseq.rds"))
```

## Alpha Diversity

```{r alpha-diversity, fig.wdth=9}
# plot_richness expects raw counts (do not rarefy/transform first)
p_alpha <- plot_richness(ps, x = "Day", measures = c("Shannon", "Simpson"), color = "When")
ggsave(filename = file.path(plot_dir, "alpha_diversity.png"), plot = p_alpha, width = 9, height = 5, dpi = 700)
```

## Ordination (NMDS, Bray-Curtis)

```{r ordination}
ps.prop <- transform_sample_counts(ps, function(x) x / sum(x))
ord.nmds.bray <- ordinate(ps.prop, method = "NMDS", distance = "bray")

p_ord <- plot_ordination(ps.prop, ord.nmds.bray, color = "When", title = "NMDS (Bray-Curtis)")
ggsave(filename = file.path(plot_dir, "ordination_nmds_bray.png"), plot = p_ord, width = 7, height = 5, dpi = 700)
```

## Abundance barplot (abundant families only)

```{r abundance-barplot, fig.width=11}
abundant <- taxa_names(ps)[taxa_sums(ps) > 500]
ps.ab <- transform_sample_counts(prune_taxa(abundant, ps), function(x) x / sum(x))

p_bar <- plot_bar(ps.ab, x = "Sample", fill = "Family") + facet_wrap(~When, scales = "free_x")
ggsave(filename = file.path(plot_dir, "abundance_plot_family.png"), plot = p_bar, width = 14, height = 10, dpi = 700)
```

## Step 11: Differential Abundance Analysis

**Goal:** find ASVs/taxa whose abundance differs between *Early* and *Late* samples. Two complementary methods will be used: - **DESeq2**: Fast, familiar, integrates directly with the `ps` object

- **ANCOM-BC2**: compositionally-aware method with bias correction. Generally more appropriate default for microbiome relative-abundance data *NB: agreement between methods is reassuring*

```{r da-subset}
# build a clean data subset - drop controls and any empty ASVs
ps <- readRDS(file.path(rds_dir, "phyloseq.rds"))

ps_da <- subset_samples(ps, !is.na(When))
ps_da <- prune_taxa(taxa_sums(ps_da) > 0, ps_da)

# prevalence filter: keep ASVs seen in >= 10% of samples - remove ultra-rare taxa that are just noise
prev <- apply(otu_table(ps_da), 2, function(x) sum(x > 0))
keep <- prev >= max(2, ceiling(0.10 * nsamples(ps_da)))
ps_da <- prune_taxa(taxa_names(ps_da)[keep], ps_da)

# fix reference level so positive results = enriched in "Late"
sample_data(ps_da)$When <- factor(sample_data(ps_da)$When, levels = c("Early", "Late"))

ps_da
```

## 11a. DESeq2

```{r da-deseq2}
library(DESeq2)

# Gaurd against empty samples / taxa left behind by the prevalence filter
ps_da <- prune_samples(sample_sums(ps_da) > 0, ps_da)
ps_da <- prune_taxa(taxa_sums(ps_da) > 0, ps_da)

dds <- phyloseq_to_deseq2(ps_da, ~ When)

dds <- estimateSizeFactors(dds, type = "poscounts")

# sanity check - no NA size factors now
stopifnot(all(!is.na(sizeFactors(dds))))

dds <- DESeq(dds, test = "Wald", fitType = "local")
res <- results(dds, cooksCutoff = FALSE)

alpha <- 0.05
sigtab <- as.data.frame(res)
sigtab <- cbind(sigtab, as(tax_table(ps_da)[rownames(sigtab), ], "matrix"))
sigtab <- sigtab[!is.na(sigtab$padj) & sigtab$padj < alpha, ]
sigtab <- sigtab[order(sigtab$log2FoldChange), ]

cat("DESeq2: significant ASVs at padj <", alpha, ":", nrow(sigtab), "\n")
head(sigtab[, c("baseMean", "log2FoldChange", "padj", "Phylum", "Family", "Genus")], 20)

write.csv(as.data.frame(res), file = file.path(table_dir, "deseq2_results.csv"), row.names = TRUE)
write.csv(sigtab, file = file.path(table_dir, "deseq2_sigtab.csv"), row.names = TRUE)

saveRDS(res, file.path(rds_dir, "deseq2_results.rds"))
saveRDS(sigtab, file.path(rds_dir, "deseq2_sigtab.rds"))
```

```{r da-deseq2-plot, fig.width=9, fig.height=6}
if (nrow(sigtab) > 0) {
  sigtab$Genus <- factor(sigtab$Genus, levels = unique(sigtab$Genus[order(sigtab$log2FoldChange)]))

p_deseq <- ggplot(sigtab, aes(x = log2FoldChange, y = Genus, color = Phylum)) + geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") + geom_point(size = 3, alpha = 0.8) + labs(title = "DESeq2: Early vs Late (positive = enriched in Late)", x = "log2 fold change", y = "Genus") + theme(axis.text.y = element_text(size = 8))

ggsave(filename = file.path(plot_dir, "da_deseq2_plot.png"), plot = p_deseq, width = 18, height = 12, dpi = 700)
}
```

## 11b. ANCOM-BC2

ANCOM-BC2 models the sampling fraction bias inherent to compositional data, which DESeq2 does not. Run at the **Genus** level here (aggregating ASVs to genus is common for DA and reduces multiple-testing burden); To test ASVs directly, set "tax_level = NULL"

```{r da-ancombc2}
library(ANCOMBC)
library(microbiome)

set.seed(123)
anc <- ancombc2(
  data = ps_da,
  tax_level = "Genus",
  fix_formula = "When",
  group = "When",
  p_adj_method = "BH",
  prv_cut = 0.10, # prevalence filter inside ANCOM-BC
  struc_zero = TRUE,
  n_cl = TAX_THREADS,
  verbose = TRUE
)

anc_res <- anc$res
# effect/q-value cols are named after the contrast, e.g. "WhenLate"
lfc_col <- grep("^lfc_When", names(anc_res), value = TRUE)
q_col <- grep("^q_When", names(anc_res), value = TRUE)
diff_col <- grep("^diff_When", names(anc_res), value = TRUE)
anc_sig <- anc_res[anc_res[[diff_col]] == TRUE, c("taxon", lfc_col, q_col)]
anc_sig <- anc_sig[order(anc_sig[[lfc_col]]), ]

cat("ANCOM-BC2: differentially abundant genera:", nrow(anc_sig), "\n")
head(anc_sig, 20)

write.csv(anc_sig, file = file.path(table_dir, "differentially abundant genera.csv"), row.names = TRUE)
saveRDS(anc, file.path(rds_dir, "ancombc2_full.rds"))
saveRDS(anc_sig, file.path(rds_dir, "ancombc2_sig.rds"))
```

```{r da-ancombc2-lot, fig.width=9, fig.height=6}
if (nrow(anc_sig) > 0) {
  anc_sig$taxon <- factor(anc_sig$taxon, levels = anc_sig$taxon[order(anc_sig[[lfc_col]])])
  
p_ancom <- ggplot(anc_sig, aes(x = .data[[lfc_col]], y = taxon)) + geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") + geom_col(fill = "steelblue", alpha = 0.85) +
    labs(title = "ANCOM-BC2: Early vs Late (positive = enriched in Late)", x = "Bias-corrected log gold change", y = "Genus")
}

ggsave(filename = file.path(plot_dir, "da_ancombc2_plot.png"), plot = p_ancom, width = 18, height = 12, dpi = 700)
```

## 11c. Cross-check the two methods

> **Interpretation.** Taxa flagged by **both** methods are your most robust hits — report those with the most confidence. Method-specific hits are worth mentioning but should be flagged as less certain. ANCOM-BC2 is generally the more defensible primary result for compositional data; DESeq2 serves as a familiar sanity check.

```{r da-compare}
# genera flagged by DESeq2 (mapped from ASV-level results)
deseq_genera <- unique(na.omit(sigtab$Genus))
ancom_genera <- if (nrow(anc_sig))
  as.character(anc_sig$taxon) else character(0)

cat("DESEq2 genera:", length(deseq_genera), 
    "| ANCOM-BC2 genera:", length(ancom_genera),
    "| overlap:", length(intersect(deseq_genera, ancom_genera)), "\n")
intersect(deseq_genera, ancom_genera)
```

# Session Info
```{r session-info}
sessionInfo()
```

# Save everything in your environment to a specific file
```{save image}
save.image(file = "16S_dada2_full_data_pipeline.RData")
```